---
title: "HUDK4051: Social Network Analysis"
author: "Gian Zlupko"
date: "1/26/2018"
output: html_document
---

In this assignment you will be generating and analyzing three social networks (also known as graphs or sociograms) based on three different measures. You will be using data from:

Representing Classroom Social Structure. Melbourne: Victoria Institute of
Secondary Education, M. Vickers and S. Chan, (1981)

Available from the Index of Complex Networks ([ICON](https://icon.colorado.edu/#!/))

The data were collected by Vickers & Chan from 29 seventh grade students in a school in Victoria, Australia. Students were asked to nominate their classmates on a number of relations including the following three "layers":  

1. Who do you get on with in the class?  
2. Who are your best friends in the class?  
3. Who would you prefer to work with?  

You have a data set for each of these questions, please complete the following tasks.

## Data Wrangling

Manipulate each of the data sets so that it is suitable for building a social network using iGraph.

*If you have not previously taken HUDK4050 please sign up for office hours before you attempt this project*

```{r}

# libraries 

library(igraph) 
library(visNetwork) 
library(tidyverse) 


# manipulate get.on.with data set 'friendly' 
friendly <- read_csv("get.on.with.csv") 

friendly_edge <- friendly %>%
  select(from, to) 
head(friendly_edge) 

#g_friendly <- graph.edgelist(as.matrix(friendly_edge), 
                             #directed = TRUE)

# add gender as a vertex attribute to get.on.with data

friendly_gender <- read_csv("get.on.with.csv") 

friendly_gender <- friendly_gender %>%
  select(from, gender.from) %>%
  rename(gender = gender.from) %>%
  rename(student = from) %>%
  distinct() 


g_friendly <- graph_from_data_frame(d = friendly_edge, 
                                    vertices = friendly_gender, 
                                    directed = TRUE) 



# best friends data set 

best_friends <- read_csv("best.friends.csv") 


best_edge <- best_friends %>%
  select(from, to) 


# load best.friends again for gender attribute 

best_vertices <- read_csv("best.friends.csv") 

best_vertices <- best_vertices %>%
  select(from, gender.from) %>%
  rename(gender = gender.from) %>%
  rename(student = from) %>%
  distinct() 


g_best <- graph_from_data_frame(d = best_edge, 
                                    vertices = best_vertices, 
                                    directed = TRUE) 





# work with data set 

work_with <- read_csv("work.with.csv") 



work_with_edge <- work_with %>%
  select(from, to) 


# load work with again for gender attribute 

work_with_vertices<- read_csv("work.with.csv") 

work_with_vertices <- work_with_vertices %>%
  select(from, gender.from) %>%
  rename(gender = gender.from) %>%
  rename(student = from) %>%
  distinct() 


g_work_with <- graph_from_data_frame(d = work_with_edge, 
                                    vertices = work_with_vertices, 
                                    directed = TRUE) 


```

## Visualize the Networks

Create a graph for each of the data sets, are the graphs directed or undirected? Visualize each of the graphs you have created and color the nodes according to gender. Save pdfs of your graphs in this directory for upload to Github.

```{r}

# visualize the get.on.with network (e.g. 'friendly') 


V(g_friendly)$color <- ifelse(
  V(g_friendly)$gender == "male", "red", "blue"
)


plot(g_friendly, edge.arrow.size = 0.2, 
     vertext.label.color = "black", main = "Friendship Network")


library(threejs) 

graphjs(g_friendly, main  = "Friendship Network", 
        showLabels = TRUE) 




# visualize the best friends network using igraph 


V(g_best)$color <- ifelse(
  V(g_best)$gender == "male", "red", "blue"
)


plot(g_best, edge.arrow.size = 0.2, 
     vertext.label.color = "black", main = "Best Friends Network")



# visualize the work preference network 


V(g_work_with)$color <- ifelse(
  V(g_work_with)$gender == "male", "red", "blue"
)


plot(g_work_with, edge.arrow.size = 0.2, 
     vertext.label.color = "black", main = "Work Preference Network")



```

## Centrality Measures

Who in the class has the highest degree centrality for each measure?
```{r}

# out degree centrality for student networks

out_degree <- function(x) { 
  
  
calculation <- degree(x, mode = c("out")) 
paste("student", which.max(calculation), "has the greatest out degree centrality", sep = " ") 
  
  }

View(degree(g_best)) 


out_degree(g_friendly) 
out_degree(g_best) 
out_degree(g_work_with) 



# in-degree centrality for the student networks 


in_degree <- function(x) { 
  
  
calculation <- degree(x, mode = c("in")) 
paste("student", which.max(calculation), "has the greatest in-degree centrality", sep = " ") 
  
  }


in_degree(g_friendly) 
in_degree(g_best) 
in_degree(g_work_with) 


# calculate greatest total degree centrality for the student networks 



total_centrality <- function(x) { 
  
  
calculation <- degree(x, mode = c("all")) 
paste("student", which.max(calculation), "has the greatest total centrality", sep = " ") 
  
  }


total_centrality(g_friendly) 
total_centrality(g_best) 
total_centrality(g_work_with) 



# IQR for total degree centrality in student networks 

summary(degree(g_friendly, mode = c("all")))

summary(degree(g_best, mode = c("all")))

summary(degree(g_work_with, mode = c("all")))


# distribution of total centrality in


friendly_centrality <- degree(g_friendly, mode = c("all")) 
hist(friendly_centrality, main = "Friendship Network Centrality Distribution",
     breaks = 15, xlab = "Centrality", ylab = "Frequency")  

best_friend_centrality <- degree(g_best, mode = c("all")) 
hist(best_friend_centrality, main = "Best Friend Network Centrality Distribution",
     breaks = 15, xlab = "Centrality", ylab = "Frequency")  


best_friend_centrality_in_coming <- degree(g_best, mode = c("in")) 
hist(best_friend_centrality_in_coming, main = "In-Coming Best Friend \n Network Centrality Distribution",
     breaks = 15, xlab = "Centrality", ylab = "Frequency")


work_with_centrality <- degree(g_work_with, mode = c("all")) 
hist(work_with_centrality, main = "Work Preferences Centrality Distribution",
     breaks = 15, xlab = "Centrality", ylab = "Frequency")  




```

Does **degree centrality** mean the same thing for each of your networks? Does it make sense for each of your networks?

Degree centrality measures the same phenomenon (number of network ties) between each network but it has a different significance/interpretation for each network. For example, in the acquaintanceship and best friend networks, degree centrality indicates the number of ties that individuals have to other actors in the network. However, an individual who has high degree centrality in the best friend network indicates that the individual is. . .. 





The centrality distribution plots indicate that the distribution of friendship centrality is more normal than work preferences. The distribution of best friend total centrality is left skewed, indicating that the majority of students in the best friend network are less central (less incoming and outcoming ties). This makes sense as we would expect best friend ties to be more exclusive. 



Who in the class has the highest closeness centrality?

```{r}


# calculate greatest closeness centrality for the student networks 


closeness_centrality <- function(x) { 
  
  
calculation <- closeness(x, mode = c("all")) 
paste("student", which.max(calculation), "has the greatest closeness centrality", sep = " ") 
  
  }

closeness_centrality(g_friendly) 
closeness_centrality(g_best) 
closeness_centrality(g_work_with) 


# betweenness centrality 



betweennes_centrality <- function(x) { 
  
  
calculation <- betweenness(x, directed = TRUE, weights = NULL) 
paste("student", which.max(calculation), "has the greatest betweenness centrality", sep = " ") 
  
  }

betweennes_centrality(g_friendly) 
betweennes_centrality(g_best) 
betweennes_centrality(g_work_with) 




```

How does **betweeness centrality** differ from degree centrality? Is one more useful than the other? Does their utility differ between your three networks?


Degree centrality is a measure of how many ties an individual has in a network. It is therefore measures an individual's centrality in a network as their connectedness to others. Individuals with high centrality are highly connected. This is contrasted with betweenness centrality, which is a separate operationalization of an individual's centarlity in a network. Betweenness centrality measures the extent to which an individual brokers a tie, or falls along a path, between other individuals in a network. This type of tie is relevant to notions of informational or resource brokerage in which an individual mediates the connection between two other individuals. Individuals with high betweenness centrality are intermediaries between many others in the network and as such, are in priviledged positions of facilitating information, resources, and processes throughout the network. 



## Simple structures

Count the number of dyads and the number and type of triads using the following commands.
```{r}


dyad_census(g_friendly) 
dyad_census(g_best) 
dyad_census(g_work_with) 

```
[Documentation](http://igraph.org/r/doc/dyad_census.html)

```{r}

triad.census(g_friendly) 
triad.census(g_best) 
triad.census(g_work_with) 



# Burt's constraint 

which.max(constraint(g_friendly, nodes = V(g_friendly), weights = NULL)) 
which.min(constraint(g_friendly, nodes = V(g_friendly), weights = NULL)) 



```
[Documentation](http://igraph.org/r/doc/triad_census.html)

Do these metrics tell you anything about the networks? If so what?

Dyads and triads are important indicators for how information flows through networks. Dyads and triads represent instances of solidified relationships between actors, which are the bases of cliques and network subgroups. Information is likely to flow within the dyadic and triadic subgroups as a function of their recipricol connectedness. The number of dyadic and triadic relationships in a network could have implications for how information flows through networks. For example, it's possible that in subgroup formations, some individuals are only connected to inividuals in that subgroup and not with other individuals in the network. This may constrain the exchange of new information throughout the network (Burt, 1992). 




## Cliques

Answer the following questions using the [clique functions](http://igraph.org/r/doc/cliques.html)

What is the size of the largest clique(s) in each of the three networks?

```{r}

# friendship network 
clique_num(g_friendly)


# best friend network 
clique_num(g_best)


# work preferences network 
clique_num(g_work_with)


```

Which nodes/vertices are in the largest cliques for the three networks? Is there much overlap?

```{r}

# friendship network 
largest_cliques(g_friendly) 
friendship_clique <- largest_cliques(g_friendly) 
friendship_clique <- data.frame(unlist(friendship_clique))
# show how many times the same individuals occur in the largest cliques;
sort(table(friendship_clique$unlist.friendship_clique.))
# many individuals appeared in four cliques


mode
# best friend network 
best_friend_clique <- largest_cliques(g_best) 
best_friend_clique <- data.frame(unlist(best_friend_clique))
sort(table(best_friend_clique$unlist.best_friend_clique.))




# work preferences network 
work_with_clique <- largest_cliques(g_work_with) 
work_with_clique <- data.frame(unlist(work_with_clique))
sort(table(work_with_clique$unlist.work_with_clique.))

```

How many **maximal cliques** are there in each of the networks?

```{r}

# 64 maximal cliques in frienship network 

friendly_max_cliques <-max_cliques(g_friendly)
length(friendly_max_cliques)


# 35 maximal cliques in frienship network 


g_best_max_cliques <-max_cliques(g_best)
length(g_best_max_cliques)



# 36 maximal cliques in frienship network 

work_with_max_cliques <-max_cliques(g_work_with)
length(work_with_max_cliques)




```

## Components & Cutpoints

Find the cutpoints (articulation points) for each of the three networks you generated. What does this tell you about the graphs? Does what you find match a visual exploration of the networks?

Student 13 is an articulation point in the best friend and work preferences networks. This means that removing student 13 would disconnect the graph into two or more parts. This means that student 13 is forming an important connection with two portions of the network that otherwise would not be connected. In Burt's terminology, we can think of student 13 as bridging a structural hole that exists in the networks - a boundary in which student 13 connects two otherwise disparate network subgroups. Interestingly, a visualization of the networks shows that student 13 is forming this essential, boundary-spanning connection for students 18 and 25 in the best friend and work with networks. 

```{r}

# friendship network 
articulation_points(g_friendly) 
plot(g_friendly) 

articulation_points(g_best) 
plot(g_best)

articulation_points(g_work_with) 
plot(g_work_with)

```



```{r}


values <- as.numeric(factor(V(g_friendly)$gender))
assortativity(g_friendly, values) 



# Calculate the observed assortativity
observed.assortativity <- assortativity(g_friendly, values)

# Calculate the assortativity of the network randomizing the gender attribute 1000 times
# tell for loop, for every i in ‘results’ calculate assortativity while randomly permuting the object values each time with sample( )

results <- vector('list', 1000)
for(i in 1:1000){
  results[[i]] <- assortativity(g_friendly, sample(values))
}

# Plot the distribution of assortativity values and add a red vertical line at the original observed value
hist(unlist(results), main = "Normal Distribution of Assortativity")  
abline(v = observed.assortativity, col = "red", lty = 3, lwd=2)


```


Community Detection 

```{r}


# friendly network 
undirected_g_friendly <- as.undirected(g_friendly, mode = c("collapse")) 


V(undirected_g_friendly)$color <- ifelse(V(undirected_g_friendly)$gender == "male", "red", "blue")


community_friendly <- fastgreedy.community(undirected_g_friendly) 

# number of communities identified

length(community_friendly) 

# 2 communities were identified by the Fast Greedy Algorithm - plot to see if those were boys and girls

plot(community_friendly, undirected_g_friendly) 

plot(community_friendly, undirected_g_friendly, edge.arrow.size = 0.2, 
     vertext.label.color = "black", main = "Subgroup Clusters \n Friendship Network") 



undirected_g_best <- as.undirected(g_best, mode = c("collapse")) 

V(undirected_g_best)$color <- ifelse(V(undirected_g_best)$gender == "male", "red", "blue")
community_best <- fastgreedy.community(undirected_g_best) 
plot(community_best, undirected_g_best) 



```





## Putting it all together

Write a narrative description of the social ties among members of this 7th grade class using the network, centrality and structural metrics you have developed. Can you provide any recommendations to the teacher of this class based on your analysis? What other information would you want to know? Can you remeber being in seventh grade, does this refelct your experience?


A few features of this 7th grade network stand out from SNA. First, visualization and community detection methods identified that there are strong, homophilic relationships within gender in the friendship network. This indicates that students primarily maintain friends with the same gender. Another finding that stands out from SNA methods was the identification of Student 13 as an articulation point in multiple networks. This signifies that Student 13 is a necessary intermediary between students 18 and 25 with the rest of the class. If the teacher wants to ensure that students 18 and 25 feel included in class activites, and avoid a scenario in which student 13's absence disconnects 18 and 25, the teacher should consider ways in which 18 and 25 can form new relationships with other students. 




## Test out your skills

Find another data set that interests you on the ([ICON](https://icon.colorado.edu/#!/)) website. Load the data into igraph, visualize the graph and use one of the metrics you have learned to characterize structures within the graph.


```{r}



# import edge list and rename columns 

law_edges <- read.table("law_edges.txt", header = F)
colnames(law_edges) <- c("layerID", "nodeID", "nodeID", "weight") 

# import nodes 

law_nodes <- read.table("law_nodes.txt", header = T) 
head(law_nodes)

# node variables: status (1 = partner, 2 = associate), gender (1 = man, 2 = woman), office (1 = Boston, 2 = Hartford, 3 = Providence), nodeSeniority = years with the firm, nodeAge, practice (1 = litigation, 2 = corporate), law school (1 = harvard, 2 = yale, 3 = uconn, 4 = other) 



# generate network object - general network (will break multiplex network into component layer networks below)

g_law <- graph_from_data_frame(d = law_edges, 
                                    vertices = law_nodes, 
                                    directed = TRUE) 


# basic plot of network 
plot(g_law) 



vertex_attr(g_law, index = V(g_law))


# centrality metrics 


law_centrality <- law_nodes %>%
  mutate(degree_in = degree(g_law, mode = c("in")), degree_out = degree(g_law, mode = c("out")), degree_total = degree(g_law, mode = c("all")), strength = strength(g_law))  %>% 
  arrange(desc(strength)) 


# View individuals in law network with greatest centrality measures 
law_centrality %>%
  select(nodeID, degree_in, degree_out, degree_total, 
         strength) 


# correlation matrix between node attributes (e.g. gender, office, seniority, etc) and centrality scores . . . significant relationships?

library(corrplot) 

law_cor <- cor(law_centrality) 
corrplot(law_cor, method = "number", 
         type = "lower") 


```



Bootstrapped T-Tests for Comparison of Degree Centrality

```{r}

library(MKinfer) 

head(law_centrality) 

litigation_strength <- law_centrality %>%
  filter(nodePractice == 2) %>%
  select(strength) 

corporate_strength <- law_centrality %>%
  filter(nodePractice == 1) %>%
  select(strength) 
head(corporate_strength) 



# compare means to test if strength is signifincantly different in one practice area

boot.t.test(litigation_strength, corporate_strength) 


female_degree_total <- law_centrality %>%
  filter(nodeGender == 2) %>%
  select(degree_total) 

male_degree_total <- law_centrality %>%
  filter(nodeGender == 1) %>%
  select(degree_total) 
head(male_degree_total) 


# total degree for males across network is statistically  larger than females. However, are there strong effect sizes for gender-based differences in the work-related ties in the network? 
boot.t.test(female_degree_total, male_degree_total, alternative = "less")



```



Multiple Regression Quadratic Assignment Procedure (MR-QAP)


```{r}

library(ergm) 


# generate layer network: advice, friendship, and co-working networks 
# law_edges (1 = advice, 2 = friendship, 3 = co-work)

View(law_edges) 

View(law_nodes) 
class(law_nodes)

head(law_edges)

colnames(law_edges) <- c("layerID", "nodeIn", "nodeOut", "weight")


# advice network 
advice_edges <- law_edges %>%
  filter(layerID == 1) 
 

g_advice <- graph_from_data_frame(d = advice_edges, 
                                    vertices = law_nodes, 
                                    directed = TRUE) 


# friendship network 

friendship_edges <- law_edges %>%
  filter(layerID == 2) 
 



g_friendship <- graph_from_data_frame(d = friendship_edges, 
                                    vertices = law_nodes, 
                                    directed = TRUE) 

# co-working network 


advice_edges <- law_edges %>%
  filter(layerID == 1) %>%
  select(nodeIn, nodeOut) 
 


friendship_edges <- law_edges %>%
  filter(layerID == 2) %>% 
  select(nodeIn, nodeOut) 


coworking_edges <- law_edges %>%
  filter(layerID == 3) %>%
  select(nodeIn, nodeOut) 

# adjacency matrices: advice, friendship and coworking 


advice_adjacency <- get.adjacency(graph.edgelist(as.matrix(advice_edges), directed = TRUE)) 

friendship_adjacency <- get.adjacency(graph.edgelist(as.matrix(friendship_edges), directed = TRUE))

coworking_adjacency <- get.adjacency(graph.edgelist(as.matrix(coworking_edges), directed=TRUE))


advice_matrix <- as.matrix(advice_adjacency)
friendship_matrix <- as.matrix(friendship_adjacency)
coworking_matrix <- as.matrix(coworking_adjacency)




# calculate the correlations between the multiple networks in the law firm:


library(sna) 
library(network)

gcor(advice_matrix, friendship_matrix, coworking_matrix)

gcor(advice_matrix,friendship_matrix)
gcor(friendship_matrix, coworking_matrix)
gcor(advice_matrix, coworking_matrix)


# test QAP


qap_test <- qaptest(list(advice_matrix, friendship_matrix, coworking_matrix), gcor, g1 = 1, g2 = 2, reps = 100)

qap_test
plot(qap_test)


advice_and_friendship_model <- netlm(advice_matrix, friendship_matrix, reps = 100)
advice_and_coworking_model <- netlm(advice_matrix, coworking_matrix, reps = 100)

# model summaries - are advice networks and co-working networks more related?
summary(advice_and_friendship_model)
summary(advice_and_coworking_model)



```


The QAP finds that advice and coworking networks share more variance than advice and friendship networks (r^2 = 0.23 and r^2 = 0.17, respectively). The differences among the shared variance is not large. An explanation for this could be that individuals at the law firm seek multiple types of advice; advice related to work and advice related to life outside of work. In general, an analysis of the correlation of each networks' adjancecy matrices showed that the advice, friendship, and coworking networks were substantially related (r = 0.4). Again, the largest correlation was found between advice and coworking (r = 0.48). One possible explanation for this is that individuals in this network often seek advice from those that they work with. The lowest correlation was between friendship and co-working (r = 0.26). One possible explanation for this is that, in general throughout the network, individuals maintain some friendships with those that they work with but they prefer to list those that they do not necessarily work with as friends. 










