---
title: "HUDK4051: Social Network Analysis"
author: "Gian Zlupko"
date: "1/26/2018"
output: html_document
---

In this assignment you will be generating and analyzing three social networks (also known as graphs or sociograms) based on three different measures. You will be using data from:

Representing Classroom Social Structure. Melbourne: Victoria Institute of
Secondary Education, M. Vickers and S. Chan, (1981)

Available from the Index of Complex Networks ([ICON](https://icon.colorado.edu/#!/))

The data were collected by Vickers & Chan from 29 seventh grade students in a school in Victoria, Australia. Students were asked to nominate their classmates on a number of relations including the following three "layers":  

1. Who do you get on with in the class?  
2. Who are your best friends in the class?  
3. Who would you prefer to work with?  

You have a data set for each of these questions, please complete the following tasks.

## Data Wrangling

Manipulate each of the data sets so that it is suitable for building a social network using iGraph.

*If you have not previously taken HUDK4050 please sign up for office hours before you attempt this project*

```{r}

# libraries 

library(igraph) 
library(visNetwork) 
library(tidyverse) 


# manipulate get.on.with data set 'friendly' 
friendly <- read_csv("get.on.with.csv") 

friendly_edge <- friendly %>%
  select(from, to) 
head(friendly_edge) 

#g_friendly <- graph.edgelist(as.matrix(friendly_edge), 
                             #directed = TRUE)

# add gender as a vertex attribute to get.on.with data

friendly_gender <- read_csv("get.on.with.csv") 

friendly_gender <- friendly_gender %>%
  select(from, gender.from) %>%
  rename(gender = gender.from) %>%
  rename(student = from) %>%
  distinct() 


g_friendly <- graph_from_data_frame(d = friendly_edge, 
                                    vertices = friendly_gender, 
                                    directed = TRUE) 



# best friends data set 

best_friends <- read_csv("best.friends.csv") 


best_edge <- best_friends %>%
  select(from, to) 


# load best.friends again for gender attribute 

best_vertices <- read_csv("best.friends.csv") 

best_vertices <- best_vertices %>%
  select(from, gender.from) %>%
  rename(gender = gender.from) %>%
  rename(student = from) %>%
  distinct() 


g_best <- graph_from_data_frame(d = best_edge, 
                                    vertices = best_vertices, 
                                    directed = TRUE) 





# work with data set 

work_with <- read_csv("work.with.csv") 



work_with_edge <- work_with %>%
  select(from, to) 


# load work with again for gender attribute 

work_with_vertices<- read_csv("work.with.csv") 

work_with_vertices <- work_with_vertices %>%
  select(from, gender.from) %>%
  rename(gender = gender.from) %>%
  rename(student = from) %>%
  distinct() 


g_work_with <- graph_from_data_frame(d = work_with_edge, 
                                    vertices = work_with_vertices, 
                                    directed = TRUE) 


```

## Visualize the Networks

Create a graph for each of the data sets, are the graphs directed or undirected? Visualize each of the graphs you have created and color the nodes according to gender. Save pdfs of your graphs in this directory for upload to Github.

```{r}

# visualize the get.on.with network (e.g. 'friendly') 


V(g_friendly)$color <- ifelse(
  V(g_friendly)$gender == "male", "red", "blue"
)


plot(g_friendly, edge.arrow.size = 0.2, 
     vertext.label.color = "black", main = "Friendship Network")


library(threejs) 

graphjs(g_friendly, main  = "Friendship Network", 
        showLabels = TRUE) 




# visualize the best friends network using igraph 


V(g_best)$color <- ifelse(
  V(g_best)$gender == "male", "red", "blue"
)


plot(g_best, edge.arrow.size = 0.2, 
     vertext.label.color = "black", main = "Best Friends Network")



# visualize the work preference network 


V(g_work_with)$color <- ifelse(
  V(g_work_with)$gender == "male", "red", "blue"
)


plot(g_work_with, edge.arrow.size = 0.2, 
     vertext.label.color = "black", main = "Work Preference Network")



```

## Centrality Measures

Who in the class has the highest degree centrality for each measure?
```{r}

# out degree centrality for student networks

out_degree <- function(x) { 
  
  
calculation <- degree(x, mode = c("out")) 
paste("student", which.max(calculation), "has the greatest out degree centrality", sep = " ") 
  
  }

View(degree(g_best)) 


out_degree(g_friendly) 
out_degree(g_best) 
out_degree(g_work_with) 



# in-degree centrality for the student networks 


in_degree <- function(x) { 
  
  
calculation <- degree(x, mode = c("in")) 
paste("student", which.max(calculation), "has the greatest in-degree centrality", sep = " ") 
  
  }


in_degree(g_friendly) 
in_degree(g_best) 
in_degree(g_work_with) 


# calculate greatest total degree centrality for the student networks 



total_centrality <- function(x) { 
  
  
calculation <- degree(x, mode = c("all")) 
paste("student", which.max(calculation), "has the greatest total centrality", sep = " ") 
  
  }


total_centrality(g_friendly) 
total_centrality(g_best) 
total_centrality(g_work_with) 



# IQR for total degree centrality in student networks 

summary(degree(g_friendly, mode = c("all")))

summary(degree(g_best, mode = c("all")))

summary(degree(g_work_with, mode = c("all")))


# distribution of total centrality in


friendly_centrality <- degree(g_friendly, mode = c("all")) 
hist(friendly_centrality, main = "Friendship Network Centrality Distribution",
     breaks = 15, xlab = "Centrality", ylab = "Frequency")  

best_friend_centrality <- degree(g_best, mode = c("all")) 
hist(best_friend_centrality, main = "Best Friend Network Centrality Distribution",
     breaks = 15, xlab = "Centrality", ylab = "Frequency")  


best_friend_centrality_in_coming <- degree(g_best, mode = c("in")) 
hist(best_friend_centrality_in_coming, main = "In-Coming Best Friend \n Network Centrality Distribution",
     breaks = 15, xlab = "Centrality", ylab = "Frequency")


work_with_centrality <- degree(g_work_with, mode = c("all")) 
hist(work_with_centrality, main = "Work Preferences Centrality Distribution",
     breaks = 15, xlab = "Centrality", ylab = "Frequency")  




```

Does **degree centrality** mean the same thing for each of your networks? Does it make sense for each of your networks?

Degree centrality measures the same phenomenon (number of network ties) between each network but it has a different significance/interpretation for each network. For example, in the acquaintanceship and best friend networks, degree centrality indicates the number of ties that individuals have to other actors in the network. However, an individual who has high degree centrality in the best friend network indicates that the individual is. . .. 





The centrality distribution plots indicate that the distribution of friendship centrality is more normal than work preferences. The distribution of best friend total centrality is left skewed, indicating that the majority of students in the best friend network are less central (less incoming and outcoming ties). This makes sense as we would expect best friend ties to be more exclusive. 



Who in the class has the highest closeness centrality?

```{r}


# calculate greatest closeness centrality for the student networks 


closeness_centrality <- function(x) { 
  
  
calculation <- closeness(x, mode = c("all")) 
paste("student", which.max(calculation), "has the greatest closeness centrality", sep = " ") 
  
  }

closeness_centrality(g_friendly) 
closeness_centrality(g_best) 
closeness_centrality(g_work_with) 


# betweenness centrality 



betweennes_centrality <- function(x) { 
  
  
calculation <- betweenness(x, directed = TRUE, weights = NULL) 
paste("student", which.max(calculation), "has the greatest betweenness centrality", sep = " ") 
  
  }

betweennes_centrality(g_friendly) 
betweennes_centrality(g_best) 
betweennes_centrality(g_work_with) 




```

How does **betweeness centrality** differ from degree centrality? Is one more useful than the other? Does their utility differ between your three networks?


Degree centrality is a measure of how many ties an individual has in a network. It is therefore measures an individual's centrality in a network as their connectedness to others. Individuals with high centrality are highly connected. This is contrasted with betweenness centrality, which is a separate operationalization of an individual's centarlity in a network. Betweenness centrality measures the extent to which an individual brokers a tie, or falls along a path, between other individuals in a network. This type of tie is relevant to notions of informational or resource brokerage in which an individual mediates the connection between two other individuals. Individuals with high betweenness centrality are intermediaries between many others in the network and as such, are in priviledged positions of facilitating information, resources, and processes throughout the network. 



## Simple structures

Count the number of dyads and the number and type of triads using the following commands.
```{r}


dyad_census(g_friendly) 
dyad_census(g_best) 
dyad_census(g_work_with) 

```
[Documentation](http://igraph.org/r/doc/dyad_census.html)

```{r}

triad.census(g_friendly) 
triad.census(g_best) 
triad.census(g_work_with) 



# Burt's constraint 

which.max(constraint(g_friendly, nodes = V(g_friendly), weights = NULL)) 
which.min(constraint(g_friendly, nodes = V(g_friendly), weights = NULL)) 



```
[Documentation](http://igraph.org/r/doc/triad_census.html)

Do these metrics tell you anything about the networks? If so what?

Dyads and triads are important indicators for how information flows through networks. Dyads and triads represent instances of solidified relationships between actors, which are the bases of cliques and network subgroups. Information is likely to flow within the dyadic and triadic subgroups as a function of their recipricol connectedness. The number of dyadic and triadic relationships in a network could have implications for how information flows through networks. For example, it's possible that in subgroup formations, some individuals are only connected to inividuals in that subgroup and not with other individuals in the network. This may constrain the exchange of new information throughout the network (Burt, 1992). 




## Cliques

Answer the following questions using the [clique functions](http://igraph.org/r/doc/cliques.html)

What is the size of the largest clique(s) in each of the three networks?

```{r}

# friendship network 
clique_num(g_friendly)


# best friend network 
clique_num(g_best)


# work preferences network 
clique_num(g_work_with)


```

Which nodes/vertices are in the largest cliques for the three networks? Is there much overlap?

```{r}

# friendship network 
largest_cliques(g_friendly) 
friendship_clique <- largest_cliques(g_friendly) 
friendship_clique <- data.frame(unlist(friendship_clique))
# show how many times the same individuals occur in the largest cliques;
sort(table(friendship_clique$unlist.friendship_clique.))
# many individuals appeared in four cliques


mode
# best friend network 
best_friend_clique <- largest_cliques(g_best) 
best_friend_clique <- data.frame(unlist(best_friend_clique))
sort(table(best_friend_clique$unlist.best_friend_clique.))




# work preferences network 
work_with_clique <- largest_cliques(g_work_with) 
work_with_clique <- data.frame(unlist(work_with_clique))
sort(table(work_with_clique$unlist.work_with_clique.))

```

How many **maximal cliques** are there in each of the networks?

```{r}

```

## Components & Cutpoints

Find the cutpoints (articulation points) for each of the three networks you generated. What does this tell you about the graphs? Does what you find match a visual exploration of the networks?

## Putting it all together

Write a narrative description of the social ties among members of this 7th grade class using the network, centrality and structural metrics you have developed. Can you provide any recommendations to the teacher of this class based on your analysis? What other information would you want to know? Can you remeber being in seventh grade, does this refelct your experience?

## Test out your skills

Find another data set that interests you on the ([ICON](https://icon.colorado.edu/#!/)) website. Load the data into igraph, visualize the graph and use one of the metrics you have learned to characterize structures within the graph.

```{r}